---
title: "pal"
subtitle: "Various helper functions that make the life of a typical R coder just a little bit easier"
author: "Salim Brüggemann"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
editor_options:
  chunk_output_type: console
---

# Packages

## Check if a package is installed

In contrast to `require`, this function checks if a package is installed without attaching its namespace if so. It is [considerably faster](https://stackoverflow.com/questions/9341635/check-for-installed-packages-before-running-install-packages/38082613#38082613) than `pkg %in% rownames(installed.packages())`.

This function [was originally proposed by Stack Overflow user Artem Klevtsov](https://stackoverflow.com/a/38082613/7196903).

```{r, is-installed}
is_installed <- function(pkg)
{
  nzchar(system.file(package = pkg))
}
```

# Miscellaneous

## Determine file path of executing script

This function is inspired by an [answer from Stack Overflow user Jerry T](https://stackoverflow.com/a/36777602/7196903).

```{r, script-path}
script_path <- function()
{
  cmd_args <- commandArgs(trailingOnly = FALSE)
  needle <- "--file="
  match <- grep(x = cmd_args,
                pattern = needle)
  
  if ( length(match) > 0 )
  {
    # Rscript
    return(normalizePath(sub(needle, "", cmd_args[match])))
    
  } else
  {
    if ( !is.null(sys.frames()[[1]]$ofile) )
    {
      # `source()`d via R console
      return(normalizePath(sys.frames()[[1]]$ofile))
      
      # RStudio Run Selection, cf. http://stackoverflow.com/a/35842176/2292993  
    } else if ( is_installed("rstudioapi") )
    {
      return(normalizePath(rstudioapi::getActiveDocumentContext()$path))

    } else rlang::abort("Couldn't determine script path!'")
  }
}
```

## Print `x` as newline-separated character vector using `cat()`

```{r, cat-lines}
cat_lines <- function(x)
{
  cat(as.character(unlist(x)),
      sep = "\n")
}
```

