---
output: pal::gitlab_document
---

# `r paste0(pal::desc_value("Package"), ": ", pal::desc_value("Title"))`

`r pal::desc_value("Description")`[^notify].


[^notify]: If you discover one or more of the features of this package in a more popular tidyverse-style package, feel free [to report it](https://gitlab.com/salim_b/r/pkgs/pal/-/issues/new) to me and I will consider removing it. I strongly support standardization efforts and the avoidance of unnecessary duplication.


## Features

...

## Details

...

## Installation
```{r, install-snippet-dev, child = pkgsnippets::file("installation-notice_dev-version_gitlab.Rmd")}
```

## Development

### Literate Programming
```{r, literate-programming-snippet, child = pkgsnippets::file("literate-programming-notice.Rmd")}
```

### Coding Style
```{r, coding-style-snippet, child = pkgsnippets::file("coding-style-notice.Rmd")}
```

## See also

- The [**xfun**](https://yihui.org/xfun/) package by Yihui Xie, a non-overlapping collection of miscellaneous R functions.
